\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{September 24}
\newcommand{\assignmentassignedate}{September 17}
\newcommand{\assignmentnumber}{Two}

\newcommand{\labyear}{2019}
\newcommand{\labday}{Tuesday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{ComputeMoonDistance}}
\newcommand{\mainprogramjava}{\lstinline{ComputeMoonDistance.java}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labtwo/ComputeMoonDistance.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running GatorGrader
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitaddmainprogram}{\command{git add src/main/java/labtwo/ComputeMoonDistance.java}}
\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labtwo/ComputeMoonDistance.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

% Enable margin notes to catch student attention

\usepackage{marginnote}
\reversemarginpar
\renewcommand*{\raggedrightmarginnote}{\centering}

% Use special fonts

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}
\usepackage{fontawesome}

% Commands for margin notes, for optional use

\newcommand{\book}{\faBook}
\newcommand{\group}{\faUsers}

% Commands for making key margin notes, fully integrated

\newcommand{\caution}[1]{\null\hfill\LARGE{\faWarning{}}\newline\scriptsize{\em{#1}}}
\newcommand{\discuss}[1]{\null\hfill\LARGE{\faCommentO{}}\newline\scriptsize{\em{#1}}}
\newcommand{\resource}[1]{\null\hfill\LARGE{\faLink{}}\newline\scriptsize{\em{#1}}}
\newcommand{\think}[1]{\null\hfill\LARGE{\faCogs{}}\newline\scriptsize{\em{#1}}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 100}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 100\\Computational Expression\\
      Fall 2019\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To learn how to use GitHub to access the files for a course assignment.
Additionally, to learn how to use your laptop's operating system and software
development tools such as a ``terminal window'' and the ``Docker Desktop''. You
will also continue to practice using Slack to support communication with the
technical leaders and the course instructor. Next, you will learn how to
implement a Java program and to write a Markdown document, also discovering how
to use the course's automated grading tool to assess your progress towards
correctly completing the project.
%
Finally, you will explore how to use a variable and perform a computation in a
Java program.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a technical
  leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

\item {\bf Store your files in GitHub}. Starting with this laboratory
  assignment, you will be responsible for storing all of your files (e.g., Java
  source code and Markdown-based writing) in a Git repository using GitHub
  Classroom. Please verify that you have saved your source code in your Git
  repository by using \command{git status} to ensure that everything is
  updated. You can see if your assignment submission meets the established
  correctness requirements by using the provided checking tools on your local
  computer and by checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  laboratory assignments will be challenging for you, each part is designed to
  give you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the student technical leaders and
  course instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to use
many of the features that GitHub provides. In particular, please make sure that
you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this assignment, you should also read
Chapters 1 and 2 in the textbook, paying particularly close attention to
Sections 1.4 and 1.5 and Sections 2.2 through 2.6. Please see the instructor if
you have questions about these reading assignments.

\section*{Accessing the Laboratory Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack workspace and find the
announcement that provides a link for it. Copy this link and paste it into your
web browser. Now, you should accept the laboratory assignment and see that
GitHub Classroom created a new GitHub repository for you to access the
assignment's starting materials and to store the completed version of your
assignment. Specifically, to access your new GitHub repository for this
assignment, please click the green ``Accept'' button and then click the link
that is prefaced with the label ``Your assignment has been created here''. If
you accepted the assignment and correctly followed these steps, you should have
created a GitHub repository with a name like
``Allegheny-Computer-Science-100-Fall-2019/computer-science-100-fall-2019-lab-2-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter into the right directory you
should now type \command{cd cs100F2019}. Next, you can type the command
\command{ls} and see that there are some files or directories inside of this
directory. By typing \command{git clone} in your terminal and then pasting in
the string that you copied from the GitHub site you will download all of the
code for this assignment. For instance, if the course instructor ran the
\command{git clone} command in the terminal, it would look like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-100-F2019/computer-science-100-fall-2019-lab-2-gkapfham.git
\end{lstlisting}

\marginnote{\discuss{Understand project layout}}[.25in] After this command
finishes, you can use \command{cd} to change into the new directory. If you want
to \step{go back} one directory from your current location, then you can type
the command \command{cd ..}. Please continue to use the \command{cd} and
\command{ls} commands to explore the files that you automatically downloaded
from GitHub. What files and directories do you see? What do you think is their
purpose? Spend some time exploring, sharing your discoveries with a neighbor and
a \mbox{technical leader}.

\section*{Performing a Computation in Java with Variables}

\begin{figure}[tb]
  \begin{verbatim}
    Gregory M. Kapfhammer Tue Sep 17 11:29:16 EDT 2019
    Distance to the moon in miles: 238855
    Number of miles in a yard: 1760
    Number of yards to the moon is: 420384800
  \end{verbatim}
\vspace*{-.25in}
\caption{Output from Running the {\tt ComputeMoonDistance} Program.}
\label{fig:output}
\end{figure}

In order view the source code of a Java program or a Markdown-based writing
assignment, you need a text editor. There are many different text editors on
your laptop and you should feel free to explore these on your own. Today, you
will view and improve a new Java program in your text editor! After you are in
your ``home base'' for this laboratory assignment in your terminal window and
text editor, please load the \mainprogramsource{} and \reflection{} into your
text editor. What do you see in these files? Please review the source code and
technical writing files, making note of the locations where you see
\program{TODO} markers for the tasks that you must complete. For instance, you
should notice that the \mainprogramsource{} requires you to write the
\command{import} statements for \program{java.util.Scanner} and
\program{java.util.Date}. Why won't the Java program compile without those lines
of source code?

\marginnote{\think{Use program variables}}[.5in] This laboratory assignment
invites you to implement a Java program that displays exactly four lines of
output. The first output line should display your name and then today's date.
You can accomplish the first task by typing \command{System.out.println("Gregory
M. Kapfhammer " + new Date());}, substituting your name for the name of the
course instructor. Your program will produce three additional lines of output
once you add three more \command{System.out.println} statements. The next three
lines will feature the output of numerical values, with the final line
outputting the result of a computation with variables. Please see the Java
source code comments for more details!

Once you have installed Docker Desktop, you can use the following
\command{docker run} command to start \command{gradle grade} as a containerized
application, using the ``DockaGator'' Docker image available on DockerHub. You
can run the following command to run the \command{gradle grade} on your project:

\begin{verbatim}
docker run --rm --name dockagator \
  -v "$(pwd)":/project \
  -v "$HOME/.dockagator":/root/.local/share \
  gatoreducator/dockagator
\end{verbatim}

The aforementioned command will use \program{"\$(pwd)"} (i.e., the current
directory) as the project directory and \program{"\$HOME/.dockagator"} as the
cached GatorGrader directory. Please note that both of these directories must
exist, although only the project directory must contain some content. Generally,
the project directory should contain the source code and technical writing for
this assignment, as provided to you through GitHub during the completion of a
previous step. Additionally, the cache directory should not contain anything
other than directories and programs created by DockaGator, thus ensuring that
they are not otherwise overwritten during the completion of the assignment. If
the above \command{docker run} command does not work correctly on the Windows
operating system, then you may need to instead run the following command to work
around limitations in the terminal window. Please recall that some Windows
terminal emulators may not correctly interpret \command{\$(pwd)} or
\command{\$HOME}, meaning that you must ``expand'' them to include the correct
path for this assignment.

\begin{verbatim}
docker run --rm --name dockagator \
  -v "$(pwd):/project" \
  -v "$HOME/.dockagator:/root/.local/share" \
  gatoreducator/dockagator
\end{verbatim}

If you want to \step{build} your program you can type the command \gradlebuild{}
in your Docker container, thereby causing the Java compiler to check your
program for errors and get it ready to run. If you see any error messages, go
back into the text editor and try to figure out what you mis-typed and fix it.
Once you have solved the problem, make a note of the error and the solution for
resolving it. Re-save your program and then build it again with the
\gradlebuild{}. If you cannot build \mainprogram{} correctly, then please talk
with a technical leader or the instructor.

\marginnote{\think{Reflect on challenges}}[.75in] When all of the errors are
eliminated, you can run your program by typing \gradlerun{} in the terminal
window---this is the ``execute'' step that will run your program and produce the
designated output. You should see your name, today's date, and three more lines
of text, as given in Figure~\ref{fig:output}. Make sure there are spaces
separating words in your output (did you forget to put a space inside the
quotation marks after your last name?). If not, then repair the program and
re-build and re-run it. Once the program runs, please reflect on this process.
What step did you find to be the most challenging? Why? You should write your
reflections in a file, called \reflection{}, that uses the Markdown writing
language. You can learn more about Markdown by viewing the aforementioned GitHub
guide. To complete this aspect of the assignment, you should write technical
content that reflects on your experiences. You should focus on building a better
understanding of each command and code segment --- if there is one that you do
not understand, please ask someone! For instance, can you explain the purpose
and behavior of these statements?

\vspace*{-.1in}
\begin{itemize}
  \setlength{\itemsep}{0pt}

  \item \program{package labtwo;}
  \item \program{import java.util.Date;}
  \item \program{int milesToMoon = 238855;}
  \item \program{int yardsPerMile = 1760;}

\end{itemize}
\vspace*{-.1in}

Since this is our second laboratory assignment and you are still learning how
to use the appropriate hardware and software, don't become frustrated if you
make a mistake. Instead, use your mistakes as an opportunity for learning both
about the necessary technology and the background and expertise of the other
students in the class, the technical leaders, and the course instructor.

\section*{Checking the Correctness of Your Program and Writing}

As verified by Checkstyle tool, the source code for the \mainprogramjava{} file
must adhere to all of the requirements in the Google Java Style Guide available
at \url{https://google.github.io/styleguide/javaguide.html}. The Markdown file
that contains your reflection must adhere to the standards described in the
Markdown Syntax Guide
\url{https://guides.github.com/features/mastering-markdown/}. Finally, your
\reflection{} file should adhere to the Markdown standards established by the
\step{Markdown linting} tool available at
\url{https://github.com/markdownlint/markdownlint/} and the writing standards
set by the \step{prose linting} tool from \url{http://proselint.com/}. Instead
of requiring you to manually check that your deliverables adhere to these
industry-accepted standards, the GatorGrader tool that you will use in this
laboratory assignment makes it easy for you to automatically check if your
submission meets the established standards for correctness. For instance,
GatorGrader will check to ensure that \mainprogram{} produces exactly four lines
of output, that you use the \command{new Date()} construct in the Java source
code, and that you correctly compute the average distance to the moon in yards.

\marginnote{\caution{Write useful commits}}[.75in] To get started with the use
of GatorGrader, type the command \gatorgraderstart{} in your Docker container.
If your laboratory work does not meet all of the assignment's requirements, then
you will see a summary of the failing checks along with a statement giving the
percentage of checks that are currently passing. If you do have mistakes in your
assignment, then you will need to review GatorGrader's output, find the mistake,
and try to fix it. Once your program is building correctly, fulfilling at least
some of the assignment's requirements, you should transfer your files to GitHub
using the \gitcommit{} and \gitpush{} commands. For example, if you want to
signal that the \mainprogramsource{} file has been changed and is ready for
transfer to GitHub you would first type \gitcommitmainprogram{} in your
terminal, followed by typing \gitpush{} and checking to see that the transfer to
GitHub is successful. If you notice that transferring your code or writing to
GitHub did not work correctly, then read the errors in your terminal window and
please try to determine why, asking a technical leader or the course instructor
for assistance, if necessary.

\marginnote{\caution{Verify all checks}}[1in] After the course instructor
enables \step{continuous integration} with a system called Travis CI, when you
use the \gitpush{} command to transfer your source code to your GitHub
repository, Travis CI will initialize a \step{build} of your assignment,
checking to see if it meets all of the requirements. If both your source code
and writing meet all of the established requirements, then you will see a green
\checkmark{} in the listing of commits in GitHub after awhile. If your
submission does not meet the requirements, a red \naughtmark{} will appear
instead. The instructor will reduce a student's grade for this assignment if the
red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit
in your GitHub repository, then you satisfied all of the main checks, thereby
allowing the course instructor to evaluate other aspects of your source code and
writing, as further described in the \step{Evaluation} section of this
assignment sheet. Unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, no late work will be considered
towards your grade for this laboratory assignment.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\begin{enumerate}

\setlength{\itemsep}{0in}

\item Stored in \reflection{}, a multiple-paragraph reflection on the commands
  that you typed in the terminal window and the Java source code that you wrote
  in your text editor. This Markdown-based document should explain the input,
  output, and behavior of each command (e.g., \command{gradle build}) and the
  challenges that you confronted when using it. For every challenge that you
  encountered, please explain your solution for it.

\item A properly documented, well-formatted, and correct version of
  \mainprogramsource{} that both meets all of the established requirements and
  produces the desired output. In particular, please recall that this Java
  program should store the result of a computation in a variable and then output
  that variable's value in the terminal window.

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the requirements.
Other factors will also influence your final grade on the assignment. In
addition to studying the efficiency and effectiveness and documentation of your
Java source code, the instructor will also evaluate the correctness of your
technical writing. If your submission receives a red \naughtmark{}, the
instructor will reduce your grade for the assignment. Finally, please remember
to read your GitHub repository's \program{README.md} file for a description of
the four grades that you will receive for this laboratory assignment.

\section*{Adhering to the Honor Code}

In adherence to the Honor Code, students should complete this assignment on an
individual basis. While it is appropriate for students in this class to have
high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem
with others and the student who produces assignments that are identical to, or
merely variations on, someone else's work. Deliverables (e.g., Java source code
or Markdown-based technical writing) that are nearly identical to the work of
others will be taken as evidence of violating the \mbox{Honor Code}. Please see
the course instructor if you have questions about this policy.

\end{document}
