\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{January 29}
\newcommand{\assignmentassignedate}{January 24}
\newcommand{\assignmentnumber}{Two}

\newcommand{\labyear}{2020}
\newcommand{\labdueday}{Wednesday}
\newcommand{\labassignday}{Friday}
\newcommand{\labtime}{10:00 am}

\newcommand{\assigneddate}{Assigned: \labassignday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{DisplayArtwork}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/practicaltwo/DisplayArtwork.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

% NOTE: Could be improved, problems due to nesting

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/practicaltwo/DisplayArtwork.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 100}
\lhead{Practical Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 100\\Computational Expression\\
      Spring 2020\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Practical \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To practice using GitHub to access the files for a practical assignment.
Additionally, to practice using the operating system on your laptop and software
development programs such as a ``terminal window'' and the ``Docker desktop''.
You will also continue to practice using Slack to support communication with the
student technical leaders and the course instructor. Next, you will learn how to
implement and run a Java program, also discovering how to use the Gradle build
tool and the course's automated grading tool to assess your progress towards
correctly completing the project. Finally, you will explore how to perform
character output and to use escaped characters.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction.
  Take notes about each step that you attempt, recording your questions and
  ideas and the challenges that you faced. If you are stuck, then please tell a
  technical leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

\item {\bf Store your files in GitHub}. As in all of your past assignments, you
  will be responsible for storing all of your files (e.g., Java source code and
  Markdown-based writing) in a Git repository using GitHub Classroom. Please
  verify that you have saved your source code in your Git repository by using
  \command{git status} to ensure that everything is updated. You can see if
  your assignment submission meets the established correctness requirements by
  using the provided checking tools on your local computer and in checking the
  commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Explore teamwork and technologies}. While certain aspects of these
  assignments will be challenging for you, each part is designed to give you the
  opportunity to learn both fundamental concepts in the field of computer
  science and explore advanced technologies that are commonly employed at a wide
  variety of companies. To explore and develop new ideas, you should regularly
  communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the student technical leader and
  course instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to
use many of the features that GitHub provides. In particular, please make sure
that you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this assignment, you should also
read Chapters 1 and 2 in the course textbook, paying particularly close
attention to Sections 1.5 and 2.1. Please see the instructor or one of the
technical leaders if you have questions on these reading assignments.

\section*{Using Escaped Characters to Display Character-Based Artwork}

To access the practical assignment, you should go into the
\channel{\#announcements} channel in our Slack workspace and find the
announcement that provides a link for it. Copy this link and paste it into your
web browser. Now, you should accept the practical assignment and see that
GitHub Classroom created a new GitHub repository for you to access the
assignment's starting materials and to store the completed version of your
assignment. Specifically, to access your new GitHub repository for this
assignment, please click the green ``Accept'' button and then click the link
that is prefaced with the label ``Your assignment has been created here''. If
you accepted the assignment and correctly followed these steps, you should have
created a GitHub repository with a name like
``Allegheny-Computer-Science-101-Spring-2020/computer-science-101-spring-2020-practical-2-gkapfham''.
Unless you provide the course instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
automatically receive a failing grade for it. Please follow the steps from a
previous assignment for finding your ``home base'' for this practical
assignment; see the instructor if you are stuck on getting started.

Now, study the documentation in the provided source code to understand the type
of output that your program should produce. Note the constraints on the
character symbols that should appear in the output and the number of possible
{\tt println} statements you can use to produce the output. That is, along with
including the characters ``{\tt @@}'' in your output, your program should use an
escaped character to create ten lines of output with only nine {\tt println}
statements. If you are not sure how to create the ten output lines with only
nine {\tt println} statements, review the textbook to learn more about the
escaped character that causes the display of a ``newline character''.

Once you have installed Docker Desktop, you can use the following
\command{docker run} command to start \command{gradle grade} as a containerized
application, using the ``DockaGator'' Docker image available on DockerHub. You
can run the following command to run the \command{gradle grade} on your project:

\vspace*{-.05in}
\begin{verbatim}
docker run --rm --name dockagator \
  -v "$(pwd)":/project \
  -v "$HOME/.dockagator":/root/.local/share \
  gatoreducator/dockagator
\end{verbatim}
\vspace*{-.05in}

Don't forget to review the previous laboratory assignment sheet to learn how to
enter the Docker container's shell and run a series of commands. Finally, if the
above \command{docker run} command does not work correctly on the Windows
operating system, then you may need to instead run the following command to work
around limitations in the terminal window called \command{cmd}:

\begin{verbatim}
docker run --rm --name dockagator \
  -v "%cd%:/project" \
  -v "%HomeDrive%%HomePath%/.dockagator:/root/.local/share" \
  gatoreducator/dockagator
\end{verbatim}

% Please recall that you may need to type the previous command differently if you
% are using Windows or using a non-standards-compliant terminal on Windows. For
% instance, you may need to place the quotation marks in a different location or
% type out the fully qualified name of the directory containing this project for
% \command{\$(pwd)} and type the fully qualified name of your home directory for
% \command{\$HOME}. Don't forget that the \program{README} in your GitHub
% repository has more tips for correctly using Docker! For example, that file
% explains how to run an ``interactive terminal'' in a Docker container.

\section*{Checking the Correctness of Your Program and Writing}

As in the past assignments, you are provided with an automated tool for checking
the quality of your source code. Please note that the practical assignments do
not require you to create technical writing as you do in the laboratory
assignments. To automatically check your Java source code you can get started
with the use of GatorGrader in Docker by typing the command \gatorgraderstart{}
after entering a container's \program{/bin/bash} prompt. If you do have mistakes
in your assignment, then you will need to review GatorGrader's output, find the
mistake, and try to fix it. Also, once your program is building correctly,
fulfilling at least some of the assignment's requirements, you should transfer
your files to GitHub using the \gitcommit{} and \gitpush{} commands. For
example, if you want to signal that the \mainprogramsource{} file has been
changed and is ready for transfer to GitHub you would first type
\gitcommitmainprogram{} in your terminal, followed by typing \gitpush{} and
checking to see that the transfer to GitHub is successful. If you notice that
transferring your Java source code to GitHub did not work correctly, then please
try to determine why, asking a technical leader or the course instructor for
assistance, if necessary.

After the course instructor enables \step{continuous integration} with a system
called Travis CI, when you use the \gitpush{} command to transfer your source
code to your GitHub repository, Travis CI will initialize a \step{build} of your
assignment, checking to see if it meets all of the requirements. If both your
source code and writing meet all of the established requirements, then you will
see a green \checkmark{} in the listing of commits in GitHub after awhile. If
your submission does not meet the requirements, a red \naughtmark{} will appear
instead. The instructor will reduce a student's grade for this assignment if the
red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit
in your GitHub repository, then you satisfied all of the baseline checks and you
will earn full credit. Unless you provide the course instructor with
documentation of the extenuating circumstances that you are facing, no late work
will be considered towards your completion grade for this practical assignment.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables. Because this is a practical
assignment, you are not required to complete any technical writing.

\vspace*{-.1in}
\begin{enumerate}

\setlength{\itemsep}{0in}

\item A properly documented, well-formatted, and correct version of
  \mainprogramsource{} that both meets all of the established requirements and,
  when run, produces the specified textual output in your terminal window.

\end{enumerate}
\vspace*{-.15in}

\section*{Evaluation of Your Practical Assignment}

Practical assignments are graded on a completion --- or ``checkmark'' --- basis. If
your GitHub repository has a \checkmark{} for the last commit before the
deadline then you will receive the highest possible grade for the assignment.
However, you will fail the assignment if you do not complete it correctly, as
evidenced by a red \naughtmark{} in your commit listing, by the set deadline for
completing the project.

\section*{Adhering to the Honor Code}

In adherence to the Honor Code, students should complete this assignment on an
individual basis. While it is appropriate for students in this class to have
high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem
with others and the student who produces assignments that are identical to, or
merely variations on, someone else's work. Deliverables (e.g., Java source code)
that are nearly identical to the work of others will be taken as evidence of
violating the \mbox{Honor Code}.

% Please see the course instructor during office hours if you have questions
% about the Honor Code policy.

\end{document}
