\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{February 27}
\newcommand{\assignmentassignedate}{February 20}
\newcommand{\assignmentnumber}{Five}

\newcommand{\labyear}{2020}
\newcommand{\labday}{Thursday}
\newcommand{\labdueday}{Thursday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{WordHide}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labfive/WordHide.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitaddmainprogram}{\command{git add src/main/java/labfive/WordHide.java}}
\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labfive/WordHide.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

% Enable margin notes to catch student attention

\usepackage{marginnote}
\reversemarginpar
\renewcommand*{\raggedrightmarginnote}{\centering}

% Commands for making key margin notes, fully integrated

\usepackage{fontawesome}
\newcommand{\caution}[1]{\null\hfill\LARGE{\faWarning{}}\newline\scriptsize{\em{#1}}}
\newcommand{\discuss}[1]{\null\hfill\LARGE{\faCommentO{}}\newline\scriptsize{\em{#1}}}
\newcommand{\resource}[1]{\null\hfill\LARGE{\faLink{}}\newline\scriptsize{\em{#1}}}
\newcommand{\think}[1]{\null\hfill\LARGE{\faCogs{}}\newline\scriptsize{\em{#1}}}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 100}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 100\\Computational Expression\\
      Spring 2020\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

In addition to enhancing the skills that you have learned in the past laboratory
assignments, the purpose of this assignment is to explore the ideas of a
``class'' and an ``object'' in the Java programming language. Specifically, you
will learn how to use the methods provided by the {\tt java.lang.String} class
to inspect and manipulate a {\tt String} object. You will then apply your
knowledge of {\tt String}s to the application domain of steganography, or the
practice of ``hiding'' messages inside of non-secret content. Finally, you will
use development tools to complete a real-world programming task, using sites
like GitHub to support your completion of the laboratory assignment.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a technical
  leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

\item {\bf Store your files in GitHub}. Starting with this laboratory
  assignment, you will be responsible for storing all of your files (e.g., Java
  source code and Markdown-based writing) in a Git repository using GitHub
  Classroom. Please verify that you have saved your source code in your Git
  repository by using \command{git status} to ensure that everything is
  updated. You can see if your assignment submission meets the established
  correctness requirements by using the provided checking tools on your local
  computer and by checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  laboratory assignments will be challenging for you, each part is designed to
  give you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the student technical leaders and
  course instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

To review what you have already learned about about variables, expressions, and
user input, please read Sections 2.1--2.6 in your textbook; pay close attention
to the {\tt Scanner} methods in Figure 2.7 and the program in Listing 2.8. To
learn more about Java classes and objects and, in particular, the methods
provided by the {\tt java.lang.String} class, please study Sections 3.1--3.2 in
the textbook. You should also study all of the slides that we discussed during
class sessions.

% Finally, don't forget to examine the ``GitHub Guides'' if you have questions
% about how to create and use a Git repository. Please see the instructor if you
% have any questions about these readings.

\section*{Accessing the Laboratory Assignment on GitHub}

To access this assignment, you should go into the \channel{\#announcements}
channel in our Slack team and find the announcement that provides a link for it.
Copy this link and paste it into your web browser. Now, you should accept the
laboratory assignment and see that GitHub Classroom created a new GitHub
repository for you to access the assignment's starting materials and to store
the completed version of your assignment. Specifically, to access your new
GitHub repository for this assignment, please click the green ``Accept'' button
and then click the link that is prefaced with the label ``Your assignment has
been created here''. If you accepted the assignment and correctly followed these
steps, you should have created a GitHub repository with a name like
``Allegheny-Computer-Science-100-Spring-2020/computer-science-100-spring-2020-lab-5-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
automatically receive a failing grade for it.

\marginnote{\caution{Clone GitHub repository}}[.75in] Before you move to the
next step of this assignment, please make sure that you read all of the content
on the web site for your new GitHub repository, paying close attention to the
technical details about the commands that you will type and the output that your
program must produce. Now you are ready to download the starting materials to
your laboratory computer. Click the ``Clone or download'' button and, after
ensuring that you have selected ``Clone with SSH'', please copy this command to
your clipboard. To enter into the right directory you should now type
\command{cd cs100S2020}. Next, you can type the command \command{ls} and see
that there are some files or directories inside of this directory. By typing
\command{git clone} in your terminal and then pasting in the string that you
copied from the GitHub site you will download all of the code for this
assignment. For instance, if the course instructor ran the \command{git clone}
command in the terminal, it would look like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-100-S2020/computer-science-100-spring-2020-lab-5-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new
directory. If you want to \step{go back} one directory from your current
location, then you can type the command \command{cd ..}. Please continue to use
the \command{cd} and \command{ls} commands to explore the files that you
automatically downloaded from GitHub. What files and directories do you see?
What do you think is \mbox{their purpose}?

% Spend some time exploring, sharing your discoveries with a neighbor and a
% \mbox{teaching assistant}.

\section*{Implementing and Evaluating a Steganography Program}

Within the field of computer security, there are many sub-fields that develop
strategies for sending, receiving, and storing secret messages. Involving the
hiding of a secret ``in plain sight'', steganography is one way in which you can
create and send a secret message. For this assignment, you will implement a
program, called {\tt WordHide.java}, that will perform these operations:

\vspace*{-.1in}

\begin{enumerate}

  \itemsep0in

  \item Prompt the user for a word that is exactly ten characters in length. For
    now, {\tt WordHide} should add to the user's input any type of ``padding
    character'' before or after a word that is less than ten characters. If the
    word is longer than ten characters, then your program should simply discard
    all of the characters after the tenth one; you may also implement better
    strategies.

  \item Since the {\tt WordHide} program must output all of its characters in a
    capitalized form, you should transform the word provided by the user so that
    it only contains upper-case letters.

  \item Finally, {\tt WordHide} should output a $20 \times 20$ ``grid'' of
    letters that contain the user's word ``hidden'' inside of it. All of the
    letters in this grid should be capitalized. You should brainstorm and
    prototype different techniques for effectively hiding the user's word in the
    grid of letters. As you implement your program, you must make decisions
    about the following matters: (i) what letters will you add to the grid to
    best hide the user's word? (ii) where and how will you place the user's word
    in the grid? (iii) what features of the Java programming language will you
    use to ensure the grid is formatted properly in the terminal window? As you
    answer these questions and finish the implementation and testing of {\tt
    WordHide.java}, it may help to consider the fact that the user's word might
    be better hidden if the grid contains both some randomly chosen letters and
    some letters found in the user's input. You should devote a considerable
    amount of time to brainstorming, prototyping, and evaluating (e.g., by
    asking a friend to try your program) different ways in which you can
    effectively hide words.

\end{enumerate}

\vspace*{-.1in}

In summary, this assignment asks you to write a program, \mainprogramsource{},
that will complete a word-hiding task. For instance, it must meet the following
requirements:

\vspace*{-.1in}
\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item Contain at least four single-line comments and two multi-line comments.
\item Include calls to at least 12 \command{println} statements.
\item Include at least one call to the \command{substring} and \command{charAt} methods.
\item Declare at least 20 \command{String} variables for the lines of output.
\item Produce a total of 24 lines of non-blank output.
\item Display the contiguous word hidden in a $20 \times 20$ grid.
\item Separate the three blocks of output with a blank horizontal line.
\item Robustly handle a variety of ``tricky'' inputs (e.g., symbols, numbers,
  and spaces).
\item Strategically hide the word in a fashion that is difficult for a human
  to detect.
\item Run in a correct and robust fashion without crashing or producing an error.

\end{enumerate}
\vspace*{-.1in}

If you want to \step{build} your program you can type the command \gradlebuild{}
in your Docker container, thereby causing the Java compiler to check your
program for errors and get it ready to run. If you get any error messages, go
back into a text editor like Atom and try to figure out what you mis-typed and
fix it. Once you have solved the problem, make a note of the error and the
solution for resolving it. Re-save your program and then build it again by
re-running the \gradlebuild{}. If you cannot build \mainprogram{} correctly,
then please talk with a technical leader or the instructor. Before you go on to
the next step of the assignment you should also ensure that you understand how
\mainprogram{} could use the following source code segment to support the
reading of values from an input file. Why is exception handling needed for this
program to perform input?

\vspace*{-.1in}
\begin{verbatim}
    try {
      wordInputsFile = new File("input/wordhide_inputs.txt");
      scanner = new Scanner(wordInputsFile);
    } catch (FileNotFoundException noFile) {
      System.out.println("Unable to locate file");
    }
\end{verbatim}
\vspace*{-.1in}

\marginnote{\think{Reflect on challenges}}[.2in] When all of the errors are
eliminated, you can run your program by typing \gradlerun{} in the Docker
container---this is the ``execute'' step that will run your program and produce
the designated output. You should see your name, today's date, and the textual
output. Make sure there are spaces separating words in your output (did you
forget to put a space inside the quotation marks after your last name?). If not,
then repair the program and re-build and re-run it. Once the program runs,
please reflect on this process. What step did you find to be the most
challenging? Why? You should write your reflections in a file, called
\reflection{}, that uses the Markdown writing language. To complete this aspect
of the assignment, you should write multiple high-quality paragraphs that report
on your experiences with the various commands and Java code segments. You should
also write about your strategy for hiding the word in the $20 \times 20$
character grid.

\section*{Checking the Correctness of Your Program and Writing}

As verified by Checkstyle, the code for the \mainprogramsource{} file must
adhere to all of the requirements in the Google Java Style Guide available at
\url{https://google.github.io/styleguide/javaguide.html}. The Markdown file that
contains your reflection must adhere to the standards described in the Markdown
Syntax Guide \url{https://guides.github.com/features/mastering-markdown/}.
Finally, your \reflection{} file should adhere to the Markdown standards
established by the \step{Markdown linting} tool available at
\url{https://github.com/markdownlint/markdownlint/} and the writing standards
set by the \step{prose linting} tool from \url{http://proselint.com/}. Instead
of requiring you to manually check your deliverables, the Gradle and GatorGrader
tools that you will use in this laboratory assignment makes it easy for you to
automatically check if your submission meets these well-established standards
for correctness.

Since this is not your first laboratory assignment, you will notice that the
provided source code does not contain all of the required comments at the top of
the Java source code file. This means that you will have to inspect the source
code from previous laboratory and practical assignments to review how to create
the comments in the \mainprogramsource{} file. Moreover, the provided source
code is missing many of the lines that are needed to pass the GatorGrader
checks. Review the requirements for this Java source code file, as outlined in
the previous section. You should also study the source code of this file to
learn more about what you need to add to it.

\marginnote{\caution{Transfer to GitHub}}[.65in] To get started with the use of
GatorGrader, type the command \gatorgraderstart{} in your terminal window. If
your laboratory work does not meet all of the assignment's requirements, then
you will see a summary of the failing checks along with a statement giving the
percentage of checks that are currently passing. If you do have mistakes in your
assignment, then you will need to review GatorGrader's output, find the mistake,
and try to fix it. Once your program is building correctly, fulfilling at least
some of the assignment's requirements, you should transfer your files to GitHub
using the \gitcommit{} and \gitpush{} commands. For example, if you want to
signal that the \mainprogramsource{} file has been changed and is ready for
transfer to GitHub you would first type \gitcommitmainprogram{} in your
terminal, followed by typing \gitpush{} and then checking to see that the
transfer to your GitHub repository is successful. Make sure that you transfer
your source code and technical writing to GitHub, as the course instructor
cannot access your deliverables unless you run the \gitpush{} command. Please
see the instructor if you cannot upload your deliverables.

\marginnote{\caution{Review all checks}}[.85in] After the course instructor
enables \step{continuous integration} with a system called Travis CI, when you
use the \gitpush{} command to transfer your source code to your GitHub
repository, Travis CI will initialize a \step{build} of your assignment,
checking to see if it meets all of the requirements. If both your source code
and writing meet all of the established requirements, then you will see a green
\checkmark{} in the listing of commits in GitHub after awhile. If your
submission does not meet the requirements, a red \naughtmark{} will appear
instead. The instructor will reduce a student's grade for this assignment if the
red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit
in your GitHub repository, then you satisfied all of the main checks, thereby
allowing the course instructor to evaluate other aspects of your source code and
writing, as further described in the \step{Evaluation} section of this
assignment sheet. Unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, no late work will be considered
towards your grade for this laboratory assignment. In conclusion, here are some
points to remember for creating a program that performs steganography:

\vspace*{-.05in}

\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item You should think carefully about how the $20 \times 20$ grid can be
  displayed using variables.

\item As in past assignments, your program only needs to have one {\tt main}
  method in one file.

\item See Figure 3.1 for a listing of some common \program{String} methods for
  use in your program.

\item Your program will alternate between creating and displaying textual
  output---this is okay!

\item Don't forget that you must hide the word contiguously in a $20 \times 20$
  grid. This means that you can divide the letters of the word and place them in
  the grid only if the letters form a ``chain'' in the grid. This requirement
  prohibits the random placement of the word's letters.

\item Don't forget to review the assignment sheets from the previous laboratory
  and practical assignments as they contain insights that will support your
  completion of this assignment.

\item Don't forget that you can leverage the source code from your previous
  course assignments as you are adding the required code segments to
  \mainprogramsource{}.

\end{enumerate}
\vspace*{-.15in}

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Stored in \reflection{}, a multiple-paragraph reflection on the commands
  that you typed in a text editor and the terminal window. This Markdown-based
  document should explain the input, output, and behavior of each command, the
  challenges that you confronted when using it, and your developed solutions.
  Using both sentences and Java source code, this document should also explain
  your strategy for hiding a word in the $20 \times 20$ character grid.

\item A complete and correct version of \mainprogramsource{} that both meets all
  of the established requirements and produces the desired textual output in the
  terminal.

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the requirements.
Other factors will also influence your final grade on the assignment. In
addition to studying the efficiency and effectiveness and documentation of your
Java source code, the instructor will also evaluate the correctness of your
technical writing. If your submission receives a red \naughtmark{}, the
instructor will reduce your grade for the assignment. Finally, please remember
to read your GitHub repository's \program{README.md} file for a description of
the four grades that you will receive for this laboratory assignment.

\section*{Adhering to the Honor Code}

In adherence to the Honor Code, students should complete this assignment on an
individual basis. While it is appropriate for students in this class to have
high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem
with others and the student who produces assignments that are identical to, or
merely variations on, someone else's work. Deliverables (e.g., Java source code
or Markdown-based technical writing) that are nearly identical to the work of
others will be taken as evidence of violating the \mbox{Honor Code}. Please see
the course instructor if you have questions about this policy.

\end{document}
