\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{April 15}
\newcommand{\assignmentassignedate}{April 6}
\newcommand{\assignmentnumber}{Eight}

\newcommand{\labyear}{2020}
\newcommand{\labday}{Monday}
\newcommand{\labdueday}{Wednesday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{CommandLineCalculator}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labeight/CommandLineCalculator.java}}

\newcommand{\secondprogram}{\lstinline{NumericalCalculator}}
\newcommand{\secondprogramsource}{\lstinline{src/main/java/labeight/NumericalCalculator.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}
\newcommand{\readme}{\lstinline{README.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitaddmainprogram}{\command{git add src/main/java/labeight/CommandLineCalculator.java}}
\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labeight/CommandLineCalculator.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}
\usepackage{fancyvrb}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 100}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 100\\Computational Expression\\
      Spring 2020\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

In this laboratory assignment, you will learn more about using iteration
constructs to perform numerical calculations, further explore the creation of
formatted output, practice calling methods in another Java class, practice using
conditional logic, and learn how to use {\tt while} loops, {\tt ArrayLists}, and
{\tt Iterator}s. You will create a program, comprised of two Java classes, that
performs calculations for the arithmetic mean and the minimum and maximum
values. You will also learn how to modify the inputs so that you can extensively
test your program and establish a confidence in its correctness.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a technical
  leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

\item {\bf Store your files in GitHub}. Starting with this laboratory
  assignment, you will be responsible for storing all of your files (e.g., Java
  source code and Markdown-based writing) in a Git repository using GitHub
  Classroom. Please verify that you have saved your source code in your Git
  repository by using \command{git status} to ensure that everything is
  updated. You can see if your assignment submission meets the established
  correctness requirements by using the provided checking tools on your local
  computer and by checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  laboratory assignments will be challenging for you, each part is designed to
  give you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the student technical leaders and
  course instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

After reviewing the GitHub materials, all of the assignment sheets for the past
laboratory and practical sessions, and the course slides and notes, you should
study Sections 5.1 through 5.8 of your textbook. As in a past assignment, you
should also review the textbook's diagrams that are in Figures 4.7 and 4.8. See
the course instructor if you have any questions about these readings.

\section*{Accessing the Laboratory Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack team and find the announcement
that provides a link for it. Copy this link and paste it into your web browser.
Now, you should accept the laboratory assignment and see that GitHub Classroom
created a new GitHub repository for you to access the assignment's starting
materials and to store the completed version of your assignment. Specifically,
to access your new GitHub repository for this assignment, please click the
green ``Accept'' button and then click the link that is prefaced with the label
``Your assignment has been created here''. If you accepted the assignment and
correctly followed these steps, you should have created a GitHub repository
with a name like
``Allegheny-Computer-Science-100-Spring-2020/computer-science-100-spring-2020-lab-8-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
will receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter into the right directory you
should now type \command{cd cs100F2020}. Next, you can type the command
\command{ls} and see that there are some files or directories inside of this
directory. By typing \command{git clone} in your terminal and then pasting in
the string that you copied from the GitHub site you will download all of the
code for this assignment. For instance, if the course instructor ran the
\command{git clone} command in the terminal, it would look like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-100-F2020/computer-science-100-spring-2020-lab-8-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new
directory. If you want to \step{go back} one directory from your current
location, then you can type the command \command{cd ..}. Please continue to use
the \command{cd} and \command{ls} commands to explore the files that you
automatically downloaded from GitHub. What files and directories do you see?
What do you think is their purpose? Spend some time exploring, sharing your
discoveries with a neighbor and a \mbox{teaching assistant}.

\begin{figure}[tb]
\begin{Verbatim}[commandchars=\\\{\}]
Gregory M. Kapfhammer Fri Apr 03 20:40:21 EDT 2020
Welcome to the Command Line Calculator!

Okay, I read in 5 values. Here they are:

25
125
19
-10
154

Nice, I found the minimum value: -10
Great, I found the maximum value: 154
Okay, I calculated the arithmetic mean as: 62.6

Thanks for using the CommandLineCalculator.
\end{Verbatim}
\vspace*{-.1in}
\caption{Sample ``{\tt CommandLineCalculator}'' output featuring numerical values read from a file.}~\label{fig:output}
\vspace*{-.1in}
\end{figure}

\section*{Extending and Implementing a Numerical Computation Engine}

You should explore your repository by using a text editor to study the source
code of the {\tt CommandLineCalculator.java} and {\tt NumericalCalculator.java}
files. What methods do these classes provide? How do they work? Can you draw a
picture that depicts the relationship between these two Java classes? Do you see
how one class will call the methods of the other? Before you start the next step
of this assignment, make sure that you understand what parts of the system are
already implemented and what you have to extend and add in order to complete the
assignment.

You will notice that this assignment organizes the methods into two separate
classes, as you have seen in past assignments and in-class exercises. In
particular, the {\tt CommandLineCalculator} provides the user interface for our
program and the {\tt NumericalCalculator} furnishes the methods that perform the
required computations. If you want to make changes to the way in which the
program accepts input or produces output, then you will need to modify the {\tt
CommandLineCalculator}. Otherwise, if you want to modify the way in which the
program performs a computation, or add a new computation, then you must make
changes to the {\tt NumericalCalculator}. Overall, these two Java classes
complete their work by following a pattern similar to that which is outlined in
Figures 4.7 and 4.8 of your textbook. Please see the instructor if you have
questions about this approach.

Additionally, it is important to note that this assignment asks you to finish
segments in both the {\tt CommandLineCalculator.java} and {\tt
NumericalCalculator.java} files. For instance, step three and steps five through
nine are not finished in the \mainprogram{}. This means that you will have to
refer to previous assignments to learn more about how to read input from a file.
Then, to perform computations you will invoke all of the methods in the
\secondprogram{}. Yet, the \secondprogram{} also contains several methods in
need of a finished implementation. Please refer to the \readme{} file in your
repository for all of the requirements associated with these files. Importantly,
you should review the partial implementation of the \program{public static
double calculateArithmeticMean(ArrayList<Integer> numberList)} method as it
provides an illustrative example of how to iterate through the
\program{numberList} in order to perform a numerical computation. To further
help you to understand how to perform both an iteration and the requested
computation, please also refer to the following source code segment for the
\program{displayNumberList} method.

\begin{verbatim}
  public static void displayNumberList(ArrayList<Integer> numberList) {
    Iterator<Integer> numberListIterator = numberList.iterator();
    // iterator through the list and display the values
    while (numberListIterator.hasNext()) {
      System.out.println(numberListIterator.next());
    }
  }
\end{verbatim}

\section*{Checking the Correctness of Your Program and Writing}

As verified by Checkstyle, the code for the \mainprogramsource{} file must
adhere to all of the requirements in the Google Java Style Guide available at
\url{https://google.github.io/styleguide/javaguide.html}. The Markdown file that
contains your reflection must adhere to the standards described in the Markdown
Syntax Guide \url{https://guides.github.com/features/mastering-markdown/}.
Finally, your \reflection{} file should adhere to the Markdown standards
established by the \step{Markdown linting} tool available at
\url{https://github.com/markdownlint/markdownlint/} and the writing standards
set by the \step{prose linting} tool from \url{http://proselint.com/}. Instead
of requiring you to manually check that your deliverables adhere to these
industry-accepted standards, the GatorGrader tool that you will use in this
laboratory assignment makes it easy for you to automatically check if your
submission meets these well-established standards for correctness. Please see
the instructor if you have questions about GatorGrader.

Since this is not your first laboratory assignment, you will notice that the
provided source code does not contain all of the required comments at the top of
the Java source code file. This means that you will have to inspect the source
code from previous laboratory and practical assignments to review how to create
the comments in the \mainprogramsource{} file. Moreover, the provided source
code is missing many of the lines that are needed to pass the GatorGrader
checks. Review the requirements for these Java code files, as outlined in the
previous section. You can study the source code of that file to learn more about
what you need to add to it. Don't forget to look in your GitHub repository to
learn about GatorGrader's many checks!

To get started with the use of GatorGrader, type the command \gatorgraderstart{}
in your terminal. If your laboratory work does not meet all of the assignment's
requirements, then you will see a summary of the failing checks along with a
statement giving the percentage of checks that are currently passing. If you do
have mistakes in your assignment, then you will need to review GatorGrader's
output, find the mistake, and try to fix it. Once your program is building
correctly, fulfilling at least some of the assignment's requirements, you should
transfer your files to GitHub using the \gitcommit{} and \gitpush{} commands.
For example, if you want to signal that the \mainprogramsource{} file has been
changed and is ready for transfer to GitHub you would first type
\gitcommitmainprogram{} in your terminal, followed by typing \gitpush{} and then
checking to see that the transfer to GitHub was successful. Make sure that you
transfer your source code and technical writing to GitHub, as the instructor
cannot access your deliverables unless you run the \gitpush{} command. Please
see the instructor if you cannot upload your deliverables.

After the course instructor enables \step{continuous integration} with a system
called Travis CI, when you use the \gitpush{} command to transfer your source
code to your GitHub repository, Travis CI will initialize a \step{build} of your
assignment, checking to see if it meets all of the requirements. If both your
source code and writing meet all of the established requirements, then you will
see a green \checkmark{} in the listing of commits in GitHub after awhile. If
your submission does not meet the requirements, a red \naughtmark{} will appear
instead. The instructor will reduce a student's grade for this assignment if the
red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit
in your GitHub repository, then you satisfied all of the main checks, thereby
allowing the course instructor to evaluate other aspects of your source code and
writing, as further described in the \step{Evaluation} section of this
assignment sheet. Unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, no late work will be considered
towards your grade for this laboratory assignment. In conclusion, here are some
points to remember for creating programs that performs computations:

\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item The provide source code is incomplete---make sure that you add all of the
  needed features.

\item You should draw a technical diagram to show the relationships between the
  Java classes.

\item Make sure that you review the textbook's diagrams that explain parameter
  passing in Java.

\item Make sure to carefully practice implementing iteration constructs and
  conditional logic.

\item Make sure to know how \program{numberList.iterator()} ``extracts'' an
  \program{Iterator} from an \program{ArrayList}.

\item As in past assignments, your program only needs to have one {\tt main}
  method in one file.

\item Your program will alternate between creating and displaying textual
  output---this is okay!

\item With that said, please carefully notice that the user input and output in
  this program is primarily in the \program{NumericalCalculator} class. Why is
  the program designed this way?

\item Your program should support reading in a variable number of data values
  from the provided text file. To ensure its correctness, make sure that you
  test the program with different inputs! For instance, you should try a
  different number of inputs and different numerical values.

\item Don't forget to review the assignment sheets from the previous laboratory
  and practical assignments as they contain insights that will support your
  completion of this assignment.

\end{enumerate}

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Stored in \reflection{}, a one-paragraph reflection on the commands that
  you typed in a text editor and the terminal window. This Markdown-based
  document should explain the input, output, and behavior of each command and
  the challenges that you confronted when using it. For every challenge that you
  encountered, please explain your solution for it. This document should also
  explain how you fixed each of the defects in the provided code.

\item A complete and correct version of \mainprogramsource{} that meets all of
  the set requirements and produces the desired textual output in the terminal.

\item A complete and correct version of \secondprogramsource{} that meets all of
  the set requirements and supports the desired textual output in the terminal.

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the requirements.
Other factors will also influence your final grade on the assignment. In
addition to studying the efficiency and effectiveness and documentation of your
Java source code, the instructor will also evaluate the correctness of your
technical writing. If your submission receives a red \naughtmark{}, the
instructor will reduce your grade for the assignment. Finally, please remember
to read your GitHub repository's \program{README.md} file for a description of
the four grades that you will receive for this laboratory assignment.

It is understood that an important part of the learning process in any course,
and particularly one in computer science, derives from thoughtful discussions
with teachers and fellow students. Such dialogue is encouraged. However, it is
necessary to distinguish carefully between the student who discusses the
principles underlying a problem with others and the student who produces
assignments that are identical to, or merely variations on, someone else's work.
While it is acceptable for students in this class to discuss their programs,
data sets, and reports with their classmates, deliverables that are nearly
identical to the work of others will be taken as evidence of violating the
\mbox{Honor Code}. You must uphold the Honor Code while completing this
assignment. Students who do not understand how to adhere to the Honor Code
should talk with the course instructor during office hours.

\end{document}
